### 2025-09-07 lecturenotes_aima331/ch1_2_aima.tex
line integrals

\begin{document}
\chapter*{Vector Calculus Continued...}
The vector integral calculus extends integrals as known from calculus to integrals over curves (``line integrals''), surfaces(``surface integrals''), and solids. The known integral from calculus is the theory of integration over coordinate lines and planes which is extended in vector calculus to general curves and surfaces in space. We shall see that these integrals have basic engineering applications in solid mechanics, in fluid flow, and in heat problems.

@@

@@

\subsection{Exercise}
\begin{enumerate}
\item Calculate work done by the given force along the given path.
  \begin{enumerate}
  \item $F=(y^3, x^3)$,  $C$ is the parabola $:y=5y^2$ from $A=(0,0)$ to $B=(2,20)$.
  \item $F=(e^x, e^y)$, $C$ is circle with center $(0,0)$ from $(1,0)$ to $(0,-1)$, clockwise.
 \item $F=(e^x, e^y, e^z)$, $C:r=(t, t^2, t^2)$ from $(0,0,0)$ to $(2,4,4)$.
  \end{enumerate}
\end{enumerate}

@@

\section{Surface Integrals}
Like line integrals, surface integrals arise in two forms. One form occurs when we integrate a scalar function over a surface, the second form is for surface integrals of vector fields.
\subsection{Surface Integrals of a vector field}
Let $F$ be a vector field in three-dimensional space with continuous components defined over a smooth surface $S$ having a chosen field of normal unit vectors $n$ orienting $S$. Then the surface integral of $F$ over $S$ is:
\begin{equation}
  \label{surf}
  \int\int_S \, F. \, \vec{d\sigma} \;\;= \;\; \int\int_S \, F. \hat{n}\, d\sigma
\end{equation}
where, $\vec{d\sigma}$ is the surface element $S$ that has direction towards its normal. \\[2mm]
The surface integral of $F$ is also called the \textbf{flux} of the vector field across the oriented surface $S$. If $F$ is a \textbf{velocity field} of a three dimensional fluid flow, then the flux $F$ across $S$ is the net rate at which fluid is crossing $S$ per unit time in the chosen positive direction $n$ defined by the orientation of $S$.

@@

  \subsubsection{Surface Integral a Non-parametrized surface}
  \begin{enumerate}
  \item For a scalar function: \\[1mm]
    $\displaystyle \int\int_S\, F\frac{|N|}{|N.k|}\,dx\,dy$
\vspace{1cm}
      \item For a vector function: \\[1mm]
    $\displaystyle \int\int_S\, F.\hat{n}\;\frac{|N|}{|N.k|}\,dx\,dy$
  \end{enumerate}

>>>>

\begin{document}
\chapter*{Vector Calculus Continued...}

The vector integral calculus extends integrals as known from calculus to integrals over curves (``line integrals''), surfaces(``surface integrals''), and solids. The known integral from calculus is the theory of integration over coordinate lines and planes which is extended in vector calculus to general curves and surfaces in space. We shall see that these integrals have basic engineering applications in solid mechanics, in fluid flow, and in heat problems.

\section{Integral of Vector Function}
The indefinite integral of $\vec{r}$ with respect to $t$ is given by:
$$\int \, \vec{r}(t)\,dt = \vec{R}(t) + C$$
If the components of $\vec{r}=f(t)\hat{i} + g(t)\hat{j}+h(t)\hat{k}$ are integrable over $[a,b]$, then so is $\vec{r}$, and the definite integral of $\vec{r}$ from $a$ to $b$ is
$$\int_{a}^{b} \; \vec{r}(t)\,dt = \int_{a}^{b}\,f(t)\,dt\,\hat{i} + \int_{a}^{b}\,g(t)\,dt\,\hat{j} + \int_{a}^{b}\,h(t)\,dt\,\hat{k}$$

@@

\subsection{Exercise}
\begin{enumerate}
\item Calculate work done by the given force along the given path.
  \begin{enumerate}
  \item $F=(y^3, x^3)$,  $C$ is the parabola $:y=5x^2$ from $A=(0,0)$ to $B=(2,20)$.
  \item $F=(e^x, e^y)$, $C$ is circle with center $(0,0)$ from $(1,0)$ to $(0,-1)$, clockwise.
  \item $F=(e^x, e^y, e^z)$, $C:r=(t, t^2, t^2)$ from $(0,0,0)$ to $(2,4,4)$.
   \item Find the work done by the force field $\vec{F} = (y-x^2)\hat{i} + (z-y^2)\hat{j} + (x-z^2)\hat{k}$ along the curve $\vec{r}(t) = t\hat{i} + t^2\hat{j} + t^3\hat{k}$, from $(0,0,0)$ to $(1,1,1)$.
  \end{enumerate}
\end{enumerate}

\begin{figure}[h]
  \centering
  \includegraphics[width=16cm, height=8cm]{circulation.png}
\end{figure}

\subsection{Flow Integrals and Circulation}
Suppose that $\vec{F}$ represents the velocity field of a fluid flowing through a region in space. Under these circumstances, the line integral along a curve in the region gives the fluid's flow along, or circulation around, the curve. In this case the line integral is called a flow integral. And if the curve end and start at the same point then it is called a circulation.

\subsection{Flux Across a Simple Closed Plane Curve}
A curve in the $xy$-plane is simple if it does not cross itself. If $C$ is a smooth simple closed curve in the domain of a continuous vector field $\vec{F}$ in the plane, and if $n$ is the outward unit normal vector on $C$, the flux of $\vec{F}$ across $C$ is $\displaystyle \int_C\; \vec{F}.\vec{n}\;ds$

@@

@@

\section{Surface Integrals}
Like line integrals, surface integrals arise in two forms. One form occurs when we integrate a scalar function over a surface, the second form is for surface integrals of vector fields. \\
To compute the mass of a surface, the flow of a liquid across a curved membrane, or the total electrical charge on a surface, we need to integrate a function over a curved surface in space. An example of this form occurs when we want to measure the net flow of a fluid across a surface submerged in the fluid.

\subsection{Surface Integrals of a vector field}
Let $F$ be a vector field in three-dimensional space with continuous components defined over a smooth surface $S$ having a chosen field of normal unit vectors $n$ orienting $S$. Then the surface integral of $F$ over $S$ is:
\begin{equation}
  \label{surf}
  \int\int_S \, F. \, \vec{d\sigma} \;\;= \;\; \int\int_S \, F. \hat{n}\, d\sigma
\end{equation}
where, $\vec{d\sigma}$ is the surface element $S$ that has direction towards its normal. \\[2mm]
The surface integral of $F$ is also called the \textbf{flux} of the vector field across the oriented surface $S$. If $F$ is a \textbf{velocity field} of a three dimensional fluid flow, then the flux $F$ across $S$ is the net rate at which fluid is crossing $S$ per unit time in the chosen positive direction $n$ defined by the orientation of $S$.

@@

  \subsubsection{Surface Integral a Non-parametrized surface}
  \begin{enumerate}
  \item For a scalar function: \\[1mm]
    $\displaystyle \int\int_S\, F\frac{|\vec{N}|}{|\vec{N}.\hat{k}|}\,dx\,dy$

      \item For a vector function: \\[1mm]
    $\displaystyle \int\int_S\, F.\hat{n}\;\frac{dx\,dy}{|\hat{n}.\hat{k}|} = \int\int_S\, F.\vec{N}\;\frac{dx\,dy}{|\vec{N}.\hat{k}|}$
  \end{enumerate}

  \begin{figure}[h]
  \centering
  \includegraphics[width=16cm, height=8cm]{parasurface.png}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=16cm, height=9cm]{nonparasurface.png}
\end{figure}
